Control Aware Radio Resource Allocation in Low Latency Wireless Control Systems

### 2019-02-12 Low_latency_control/response/response_main.tex

\end{itemize}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%   R   E   V   I   E   W   E   R    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\reviewer

We thank the reviewer for the comments, feedback, and questions. The reviewer raised concerns about the placement of the work within the URLL control and communication framework and literature. The reviewer also brings up points of clarification regarding the problem formulation. Each concern is addressed individually below.

\issue{ The authors need to recast the framework within ultra-reliable and low-latency control and communication.}
\answer{We thank the reviewer for this important comment and suggestion. We have revised the introduction of the manuscript substantially to contextualize the work being done here within the framework of URLLC. In doing so, we have added the following important references on the topic to the disucssion in the introduction.
\begin{itemize}
\item M. Weiner, M. Jorgovanovic, A. Sahai, and B. Nikolie, “Design of a
low-latency, high-reliability wireless communication system for control
applications,” in Communications (ICC), 2014 IEEE International
Conference on. IEEE, 2014, pp. 3829–3835.
\item P. Popovski, J. J. Nielsen, C. Stefanovic, E. de Carvalho, E. Strom,
K. F. Trillingsgaard, A.-S. Bana, D. M. Kim, R. Kotaba, J. Park et al.,
“Wireless access for ultra-reliable low-latency communication: Principles
and building blocks,” IEEE Network, vol. 32, no. 2, pp. 16–23, 2018.
\item M. Bennis, M. Debbah, and H. V. Poor, “Ultra-reliable and lowlatency
wireless communication: Tail, risk and scale,” arXiv preprint
arXiv:1801.01270, 2018.
\item V. N. Swamy, S. Suri, P. Rigge, M. Weiner, G. Ranade, A. Sahai, and
B. Nikolic, “Cooperative communication for high-reliability low-latency
wireless control,” in Communications (ICC), 2015 IEEE International
Conference on. IEEE, 2015, pp. 4380–4386.
\item J. J. Nielsen, R. Liu, and P. Popovski, “Ultra-reliable low latency communication
using interface diversity,” IEEE Transactions on Communications,
vol. 66, no. 3, pp. 1322–1334, 2018.
\end{itemize}
In addition to this more comprehensive review of URLLC techniques, we have the introduction to highlight the fact that our control-aware scheduling approach can be seen as a means of defining an \emph{adaptive}-reliability framework for achieving low-latency communications, in contrast to fixed ultra reliability targets. This adaptive reliability framework considers the particular control system dynamics to determine the reliability necessary to keep all plants in good conditions. Our numerical simulations show how this alternative framework can improve the scale of low-latency wireless control systems by loosening reliability requirements where applicable. 
}

@@

\issue{ It was unclear whether the focus is on communication for a given control or the other way around. Taht is currently it looks like the solution is based on sequential design.}
\answer{We thank the reviewer for brining up this point of clarification. The focus of this work is designing a scheduling protocol for a given system and control. As such, the control is assumed to be given, or chosen agnostic to the communication, while the communication is designed relative to both the system dynamics and choice of control. We have added a sentence to Section II to clarify this point. }

\issue{ The article would improve a lot if specific use cases are defined.}
\answer{\red{To be written}}

\issue{ Linear control was adopted (for tractability) but clearly the nonlinearity is needed}
\answer{The reviewer brings up a reasonable point about the formulation of the wireless control problem. Indeed, many practical systems feature both nonlinear dynamics and potentially non-linear controls. As stated, the proposed approach focuses on linear systems and controls so that we may derive analytic expressions for the control-aware adaptive packet delivery rates. In practice, nonlinear systems may be linearized for the purposes of obtaining such packet delivery rate targets in (23) and performing the proposed CALLS method as written. \blue{To demonstrate how this may be done, in the revised manuscript we have included additional simulations in Section VI in which we control a non-linear system using the CALLS method to schedule plant communications. In these simulations, while the actual dynamics and control being applied are nonlinear in nature, a linearization of the dynamics is used solely to derive the necessary PDR targets as in expression (23). Here, we demonstrate that the proposed method is applicable even in nonlinear systems.}   }

\issue{Reliability as not addressed however it is intertwined with latency}
\answer{This is a good point raised by the reviewer that deserves clarification. Reliability is a key component of a practical system that must be carefully managed in low latency systems. In our work, rather than aim for traditional, communication-based reliability targets, such as $0.999$ packet delivery rates, we consider the fact that the observed performance of a control system is more closely related to a \emph{control}-based reliability metric, such as stability or distance from origin. As discussed in Section III and IV, by defining reliability targets with respect to the Lyapunov cost function rather than with respect to a fixed packet delivery rate target, we are able to derive \emph{adaptive} packet delivery rate requirements that change with the current performance of each control system. When a control system is in a safe or stable range, it can remain in good performance with smaller packet delivery rates than a control system in a less desirable region. Thus, the proposed CALLS method aims to achieve reliability of control system performance, in other words an \emph{adaptive reliability}, in contrast to a standard high reliability framework. Such adaptive rates are critical in enabling low-latency transmissions in large-scale wireless settings, as observed our simulations in Section V. We point out that the proposed control-adaptive framework is thus presented as an alternative to high reliability scheduling schemes in settings in which the control systems are well-managed and defined, such as in the case of industrial controls. We have added a discussion of this to the revised manuscript.}

\issue{Authors did not cite recent and fundamental articles within URLLC:}
\answer{We thank the reviewer for pointing out these references. They have been added to the revised manuscript, as well as  included in an added discussion of URLLC in the introduction (Section I). The references added are listed in a previous response to this reviewer. }

### 2019-02-12 Low_latency_control/revised/simulations.tex
Add files via upload

In this section, we simulate the implementation of both the control-aware CALLS method and a standard ``control-agnostic'' scheduling methods for various low-latency control systems over a simulated wireless channel. We point out the low-latency based scheduling/assignment approaches of both methods being compared are identical, with the distinguishing features being the dynamic control-aware packet delivery rates incorporated in the CALLS method. In doing so, we may analyze the performance of the control-aware design outlined in the previous section relative to a standard latency-aware approach in terms of, e.g., number of users supported with fixed latency threshold or best latency achieved with fixed number of users. As we are interested primarily in low latency settings that tightly restrict the communication resources, we consider two standard control systems whose rapidly changing state requires high sampling rates, and consequently a communication latency on the order of milliseconds. The parameters for the simulation setup are provided in Table \ref{tab_simulation}. The packet delivery rate function $q(\bbh,\mu,\bbsigma)$ is computed using the standard AWGN noise curves for wireless channels. The transmission time $\tau(\mu,\bbsigma)$ is computed in the simulations using the associated data rates of an MCS in Table \ref{tab_mcs} for a 100 byte packet and overhead (e.g. TFs) of the 802.11ax specifications. The latency overhead for this setting amounts to approximately $\tau_0 \approx 100 \mu$s.
%
\begin{table}
\begin{tabular}{ l | l } \hline
  Channel model & IEEE Model E (indoor) \cite{liu2014ieee} \\
  Sensor to AP distances & Random (1 to 50 meters)\\
  Transmit power & 23 dbm \\
  Channel bandwidth & 20 MHz \\
  RU sizes & 2, 4, 8, 20 MHz \\
  \# of antennas at AP & 2 \\
  \# of antennas at sensors & 1 \\
  MCS options & See Table \ref{tab_mcs} \\
  State sampling period & 10 ms \\
  \hline
\end{tabular}
\caption{Simulation setting parameters.}
\label{tab_simulation}
\end{table}
%

@@

\begin{figure}
\input{images/pendulum.tex}
\caption{Inverted pendulum-cart system $i$. The state $\bbx_{i,k} = [x_{i,k}, \dot{x}_{i,k}, \theta_{i,k}, \dot{\theta}_{i,k}]$ contains specifies angle $ \theta_{i,k}$ of the pendulum to the vertical, while the input $u_{i,k}$ reflects a horizontal force on the cart.}
\label{fig_inverted_pendulum}
\end{figure}

@@

%%%%
\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{images/ip_dist.eps}
\caption{Average pendulum distance to center vertical for $m=25$ devices using (top) CALLS and (bottom) fixed-PDR scheduling with $\tau_{\max}= 1$ ms latency threshold. The proposed control aware scheme keeps all pendulums close to the vertical, while fixed-PDR scheduling cannot.}
\label{fig_ip_dist}
\end{figure}
%%

@@

%%%%
\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{images/bbb_dist.eps}
\caption{Average ball distance to center for $m=50$ devices using (top) CALLS and (bottom) fixed-PDR scheduling with $\tau_{\max} =1$ ms latency threshold. The proposed control aware scheme keeps all balancing balls close to center, while fixed-PDR scheduling cannot.}
\label{fig_bbb_dist}
\end{figure}
%%

%%%%
\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{images/bbb_psr.eps}
\caption{Histogram of achieved PDRs in $m=50$ balancing board systems (top) CALLS and (bottom) fixed-PDR scheduling with $\tau_{\max} =1$ ms latency threshold. The proposed control aware scheme achieves similar PDRs for all devices, while the control agnostic scheduling results in large variation in packet delivery.}
\label{fig_bbb_psr}
\end{figure}
%%

>>>>

In this section, we simulate the implementation of both the control-aware CALLS method and a standard ``control-agnostic'' scheduling methods for various low-latency control systems over a simulated wireless channel. We point out the low-latency based scheduling/assignment approaches of both methods being compared are identical, with the distinguishing features being the dynamic control-aware packet delivery rates incorporated in the CALLS method. In doing so, we may analyze the performance of the control-aware design outlined in the previous section relative to a standard latency-aware approach in terms of, e.g., number of users supported with fixed latency threshold or best latency achieved with fixed number of users. As we are interested primarily in low latency settings that tightly restrict the communication resources, we consider two standard control systems whose rapidly changing state requires high sampling rates, and consequently a communication latency on the order of milliseconds. The parameters for the simulation setup are provided in Table \ref{tab_simulation}. \red{The wireless fading channel modeled using  IEEE Indoor Channel Model E, while the packet delivery rate function $q(\bbh,\mu,\bbsigma)$ is computed using the corresponding SNR values via an AWGN noise curve.} The transmission time $\tau(\mu,\bbsigma)$ is computed in the simulations using the associated data rates of an MCS in Table \ref{tab_mcs} for a 100 byte packet and overhead (e.g. TFs) of the 802.11ax specifications. The latency overhead for this setting amounts to approximately $\tau_0 \approx 100 \mu$s.
%
\begin{table}
\begin{tabular}{ l | l } \hline
  Channel model & IEEE Model E (indoor) \cite{liu2014ieee} \\
  Sensor to AP distances & Random (1 to 50 meters)\\
  Transmit power & 23 dbm \\
  Channel bandwidth & 20 MHz \\
  RU sizes & 2, 4, 8, 20 MHz \\
  \# of antennas at AP & 2 \\
  \# of antennas at sensors & 1 \\
  MCS options & See Table \ref{tab_mcs} \\
  State sampling period & 10 ms \\
  \hline
\end{tabular}
\caption{Simulation setting parameters.}
\label{tab_simulation}
\end{table}
%

@@

\begin{figure}
\input{../images/pendulum.tex}
\caption{Inverted pendulum-cart system $i$. The state $\bbx_{i,k} = [x_{i,k}, \dot{x}_{i,k}, \theta_{i,k}, \dot{\theta}_{i,k}]$ contains specifies angle $ \theta_{i,k}$ of the pendulum to the vertical, while the input $u_{i,k}$ reflects a horizontal force on the cart.}
\label{fig_inverted_pendulum}
\end{figure}

@@

%%%%
\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{../images/ip_dist.eps}
\caption{Average pendulum distance to center vertical for $m=25$ devices using (top) CALLS and (bottom) fixed-PDR scheduling with $\tau_{\max}= 1$ ms latency threshold. The proposed control aware scheme keeps all pendulums close to the vertical, while fixed-PDR scheduling cannot.}
\label{fig_ip_dist}
\end{figure}
%%

@@

%%%%
\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{../images/bbb_dist.eps}
\caption{Average ball distance to center for $m=50$ devices using (top) CALLS and (bottom) fixed-PDR scheduling with $\tau_{\max} =1$ ms latency threshold. The proposed control aware scheme keeps all balancing balls close to center, while fixed-PDR scheduling cannot.}
\label{fig_bbb_dist}
\end{figure}
%%

%%%%
\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{../images/bbb_psr.eps}
\caption{Histogram of achieved PDRs in $m=50$ balancing board systems (top) CALLS and (bottom) fixed-PDR scheduling with $\tau_{\max} =1$ ms latency threshold. The proposed control aware scheme achieves similar PDRs for all devices, while the control agnostic scheduling results in large variation in packet delivery.}
\label{fig_bbb_psr}
\end{figure}
%%
